---
title: "Data_exploration"
author: "Adam Polinko"
date: "November 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(tidyverse)
library(openxlsx)
```


## Read the data

Write a clean names function
```{r}
clean_names <- function(.data, unique = FALSE) {
  n <- if (is.data.frame(.data)) colnames(.data) else .data

  n <- gsub("%+", "_pct_", n)
  n <- gsub("\\$+", "_dollars_", n)
  n <- gsub("\\++", "_plus_", n)
  n <- gsub("-+", "_minus_", n)
  n <- gsub("\\*+", "_star_", n)
  n <- gsub("#+", "_cnt_", n)
  n <- gsub("&+", "_and_", n)
  n <- gsub("@+", "_at_", n)

  n <- gsub("[^a-zA-Z0-9_]+", "_", n)
  n <- gsub("([A-Z][a-z])", "_\\1", n)
  n <- tolower(trimws(n))
  
  n <- gsub("(^_+|_+$)", "", n)
  
  n <- gsub("_+", "_", n)
  
  if (unique) n <- make.unique(n, sep = "_")
  
  if (is.data.frame(.data)) {
    colnames(.data) <- n
    .data
  } else {
    n
  }
}
```


```{r}
regen_data <- read_csv("regen_data.csv") %>% 
  clean_names() %>% 
  separate(gap_no, c("gap","plot"))
```

##Prep the data

Pull out 0 plots

```{r}
zero_plots <- regen_data %>% 
  filter(spp == "0")

```


Now find germinates
```{r}
regen_data <- regen_data %>% 
  mutate(germ_id = ifelse(height == growth, "germinate", "older")) 


germinate_id <- regen_data %>% 
  filter(germ_id == "germinate")

```



Now we need to find plots with only germinates. First, I will create a dataframe with plots that have germinates.
```{r}
plots_with_germinates <- regen_data %>% filter(plot %in% germinate_id$plot)

```

Now find plots with older trees as well as germinates
```{r}
plots_with_germinates_older <- plots_with_germinates %>% filter(germ_id == "older")

```

and remove those plots from the plots with germinates dataframe
```{r}
plots_with_germinates_no_old <- plots_with_germinates %>% filter(!plot %in% plots_with_germinates_older$plot)
```

Now we have a dataframe with plots with only germinates. There are several At in the plots that have NA values. We won't worry about them for now.

```{r}
plots_with_germinates_no_old <- plots_with_germinates_no_old %>% 
  mutate_at(.vars = vars( height, growth, diameter, width, damage, 
vigour), ~replace(., nchar(.) >= 1, NA)) %>% 
  mutate(spp = ifelse(nchar(spp) >= 1, "0" , spp)) %>% 
  select(-germ_id) %>% 
  distinct()
```


Now combine the zero plots in
```{r}
zero_plots2<- bind_rows(zero_plots, plots_with_germinates_no_old)
```

Now let's find advanced regeneration
```{r}
advanced_regen <- regen_data %>% 
  filter(regn_type != "N")

advanced_regen2 <- regen_data %>% 
  filter(plot %in% advanced_regen$plot)

advanced_regen3 <- advanced_regen2 %>% 
  filter(regn_type == "N")

advanced_regen4 <- advanced_regen %>% 
  filter(!plot %in% advanced_regen3$plot)

advanced_regen_final <- advanced_regen4 %>% 
    mutate_at(.vars = vars( height, growth, diameter, width, damage, 
vigour), ~replace(., nchar(.) >= 1, NA)) %>% 
  mutate(spp = ifelse(nchar(spp) >= 1, "0" , spp)) %>% 
  select(-germ_id) %>% 
  distinct()



```

Now add this back in 
```{r}
zero_plots_final <- bind_rows(zero_plots2, advanced_regen_final)
```

So now we count douglas-fir
```{r}
regen_data_ready_count <- regen_data %>% 
  filter(!plot %in% zero_plots_final$plot) %>% 
  filter(spp == "Fd", regn_type == "N") %>% 
  group_by(gap, plot) %>% 
  summarise(count = n()) 


```


NOw we create a dataframe to merge into zeros
```{r}
zero_counts <- zero_plots_final %>% 
  select(gap, plot) %>% 
  mutate(count = 0)
```


bind the two dataframes
```{r}
regen_by_plot <- bind_rows(regen_data_ready_count, zero_counts)
```


add TPH
```{r}
regen_by_plot <- regen_by_plot %>% 
  mutate(tph = count * 1000)
```

Now let's find density by gap
```{r}
regen_by_gap <- regen_by_plot %>% 
  group_by(gap) %>% 
  summarise(mean_density = mean(tph),
            sem_density = sd(tph)/sqrt(n()))
```

Now to obtain scatter plots for the 4Ring and 5Ring LAI we must upload the PASG data 

First we have to clean names 
```{r}
clean_names <- function(.data, unique = FALSE) {
  n <- if (is.data.frame(.data)) colnames(.data) else .data

  n <- gsub("%+", "_pct_", n)
  n <- gsub("\\$+", "_dollars_", n)
  n <- gsub("\\++", "_plus_", n)
  n <- gsub("-+", "_minus_", n)
  n <- gsub("\\*+", "_star_", n)
  n <- gsub("#+", "_cnt_", n)
  n <- gsub("&+", "_and_", n)
  n <- gsub("@+", "_at_", n)

  n <- gsub("[^a-zA-Z0-9_]+", "_", n)
  n <- gsub("([A-Z][a-z])", "_\\1", n)
  n <- tolower(trimws(n))
  
  n <- gsub("(^_+|_+$)", "", n)
  
  n <- gsub("_+", "_", n)
  
  if (unique) n <- make.unique(n, sep = "_")
  
  if (is.data.frame(.data)) {
    colnames(.data) <- n
    .data
  } else {
    n
  }
}
```

Loading regen_data1 with names cleaned and gap and plot numbers seperated
```{r}
regen_data1N <- read.csv('regen_data1.csv') %>% 
 clean_names() %>% 
  separate(gap_no, c("gap","plot")) 
colnames(regen_data1N)
```

Now we obtain informatiion on the LAI  
```{r}
regen_LAI <- regen_data1N%>%
  group_by(gap, plot, gap_type) %>% 
  select(lai_4_ring,lai_5_ring)
view(regen_LAI)
```

Now we get the LAI 4&5 for salavage plots
```{r}
regen_LAIS <- regen_LAI %>%
  filter(gap_type =="S")
```

Now we get the LAI 4&5 for natural plots
```{r}
regen_LAIN <- regen_LAI %>%
  filter(gap_type =="N")
```

NOw we join regen plot data and LAIS (salvage) data

```{r}
S_regen_LAI <- regen_by_plot %>% left_join(regen_LAIS) %>% 
  filter (gap_type =="S")

```

Now we obtain regen plot data and LAIN (natuaral) data

```{r}
N_regen_LAI <-  regen_by_plot %>% left_join(regen_LAIN) %>% 
  filter(gap_type =="N")

```

Now we plot scatter plots for both LAI Salavage

```{r}
SScplot_LAI4R <- plot(lai_4_ring ~ tph, data = S_regen_LAI)
SScplot_LAI4R1 <- plot (tph ~ lai_4_ring, data = S_regen_LAI) 
SScplot_LAI5R <- plot (lai_5_ring ~ tph, data = S_regen_LAI)
```

Now we plot scatter plots for both LAI Natural

```{r}
NScplot_4LAI <- plot(lai_4_ring ~tph, data= N_regen_LAI)
NScplot_5LAI <- plot(lai_5_ring ~tph, data= N_regen_LAI)
```

