---
title: "Data_exploration"
author: "Emmanuel"
date: "November 22, 2019"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(tidyverse)
library(openxlsx)
library(circular)
library(car)
```


## Read the data

Write a clean names function
```{r}
clean_names <- function(.data, unique = FALSE) {
  n <- if (is.data.frame(.data)) colnames(.data) else .data

  n <- gsub("%+", "_pct_", n)
  n <- gsub("\\$+", "_dollars_", n)
  n <- gsub("\\++", "_plus_", n)
  n <- gsub("-+", "_minus_", n)
  n <- gsub("\\*+", "_star_", n)
  n <- gsub("#+", "_cnt_", n)
  n <- gsub("&+", "_and_", n)
  n <- gsub("@+", "_at_", n)

  n <- gsub("[^a-zA-Z0-9_]+", "_", n)
  n <- gsub("([A-Z][a-z])", "_\\1", n)
  n <- tolower(trimws(n))
  
  n <- gsub("(^_+|_+$)", "", n)
  
  n <- gsub("_+", "_", n)
  
  if (unique) n <- make.unique(n, sep = "_")
  
  if (is.data.frame(.data)) {
    colnames(.data) <- n
    .data
  } else {
    n
  }
}
```


```{r}
regen_data <- read_csv("regen_data.csv") %>% 
  clean_names() %>% 
  separate(gap_no, c("gap","plot")) 
```

##Prep the data

Pull out 0 plots

```{r}
zero_plots <- regen_data %>% 
  filter(spp == "0")

```


Now find germinates
```{r}
regen_data <- regen_data %>% 
  mutate(germ_id = ifelse(height == growth, "germinate", "older")) 


germinate_id <- regen_data %>% 
  filter(germ_id == "germinate")

```

Now we need to find plots with only germinates. First, I will create a dataframe with plots that have germinates.
```{r}
plots_with_germinates <- regen_data %>% filter(plot %in% germinate_id$plot)

```

Now find plots with older trees as well as germinates
```{r}
plots_with_germinates_older <- plots_with_germinates %>% filter(germ_id == "older")

```

and remove those plots from the plots with germinates dataframe
```{r}
plots_with_germinates_no_old <- plots_with_germinates %>% filter(!plot %in% plots_with_germinates_older$plot)
```

Now we have a dataframe with plots with only germinates. There are several At in the plots that have NA values. We won't worry about them for now.

```{r}
plots_with_germinates_no_old <- plots_with_germinates_no_old %>% 
  mutate_at(.vars = vars( height, growth, diameter, width, damage, 
vigour), ~replace(., nchar(.) >= 1, NA)) %>% 
  mutate(spp = ifelse(nchar(spp) >= 1, "0" , spp)) %>% 
  select(-germ_id) %>% 
  distinct()
```


Now combine the zero plots in
```{r}
zero_plots2<- bind_rows(zero_plots, plots_with_germinates_no_old)
```

Now let's find advanced regeneration
```{r}
advanced_regen <- regen_data %>% 
  filter(regn_type != "N")

advanced_regen2 <- regen_data %>% 
  filter(plot %in% advanced_regen$plot)

advanced_regen3 <- advanced_regen2 %>% 
  filter(regn_type == "N")

advanced_regen4 <- advanced_regen %>% 
  filter(!plot %in% advanced_regen3$plot)

advanced_regen_final <- advanced_regen4 %>% 
    mutate_at(.vars = vars( height, growth, diameter, width, damage, 
vigour), ~replace(., nchar(.) >= 1, NA)) %>% 
  mutate(spp = ifelse(nchar(spp) >= 1, "0" , spp)) %>% 
  select(-germ_id) %>% 
  distinct()

```

Now add this back in 
```{r}
zero_plots_final <- bind_rows(zero_plots2, advanced_regen_final)
```

So now we count douglas-fir
```{r}
regen_data_ready_count <- regen_data %>% 
  filter(!plot %in% zero_plots_final$plot) %>% 
  filter(spp == "Fd", regn_type == "N") %>% 
  group_by(gap, plot) %>% 
  summarise(count = n()) 


```


NOw we create a dataframe to merge into zeros
```{r}
zero_counts <- zero_plots_final %>% 
  select(gap, plot) %>% 
  mutate(count = 0)
```


bind the two dataframes
```{r}
regen_by_plot <- bind_rows(regen_data_ready_count, zero_counts)
```


add TPH
```{r}
regen_by_plot <- regen_by_plot %>% 
  mutate(tph = count * 1000)
```

Now let's find density by gap
```{r}
regen_by_gap <- regen_by_plot %>% 
  group_by(gap) %>% 
  summarise(mean_density = mean(tph),
            sem_density = sd(tph)/sqrt(n()))
```

Now to obtain scatter plots for the 4Ring and 5Ring LAI we must upload the PASG data 

First we have to clean names 
```{r}
clean_names <- function(.data, unique = FALSE) {
  n <- if (is.data.frame(.data)) colnames(.data) else .data

  n <- gsub("%+", "_pct_", n)
  n <- gsub("\\$+", "_dollars_", n)
  n <- gsub("\\++", "_plus_", n)
  n <- gsub("-+", "_minus_", n)
  n <- gsub("\\*+", "_star_", n)
  n <- gsub("#+", "_cnt_", n)
  n <- gsub("&+", "_and_", n)
  n <- gsub("@+", "_at_", n)

  n <- gsub("[^a-zA-Z0-9_]+", "_", n)
  n <- gsub("([A-Z][a-z])", "_\\1", n)
  n <- tolower(trimws(n))
  
  n <- gsub("(^_+|_+$)", "", n)
  
  n <- gsub("_+", "_", n)
  
  if (unique) n <- make.unique(n, sep = "_")
  
  if (is.data.frame(.data)) {
    colnames(.data) <- n
    .data
  } else {
    n
  }
}
```

Loading regen_data1 with names cleaned and gap and plot numbers seperated
```{r}
regen_data1N <- read.csv('regen_data1.csv') %>% 
 clean_names() %>% 
  separate(gap_no, c("gap","plot")) 
colnames(regen_data1N)
```

Now we obtain informatiion on the LAI  
```{r}
regen_LAI <- regen_data1N%>%
  group_by(gap, plot, gap_type) %>% 
  select(lai_4_ring,lai_5_ring)
view(regen_LAI)
```

Now we get the LAI 4&5 for salavage plots
```{r}
regen_LAIS <- regen_LAI %>%
  filter(gap_type =="S")
```

Now we get the LAI 4&5 for natural plots
```{r}
regen_LAIN <- regen_LAI %>%
  filter(gap_type =="N")
```

NOw we join regen plot data and LAIS (salvage) data

```{r}
S_regen_LAI <- regen_by_plot %>% left_join(regen_LAIS) %>% 
  filter (gap_type =="S")

```

Now we obtain regen plot data and LAIN (natuaral) data

```{r}
N_regen_LAI <-  regen_by_plot %>% left_join(regen_LAIN) %>% 
  filter(gap_type =="N")

```

Now we plot scatter plots for both LAI Salavage

```{r}
SScplot_LAI4R <- plot(lai_4_ring ~ tph, data = S_regen_LAI)
SScplot_LAI4R1 <- plot (tph ~ lai_4_ring, data = S_regen_LAI) 
SScplot_LAI5R <- plot (lai_5_ring ~ tph, data = S_regen_LAI)
```

Now we plot scatter plots for both LAI Natural

```{r}
NScplot_4LAI <- plot(lai_4_ring ~tph, data= N_regen_LAI)
NScplot_5LAI <- plot(lai_5_ring ~tph, data= N_regen_LAI)
```

Lets fix the germinants by taking them out
```{r}
Zeroplots_germinants <- plots_with_germinates %>% 
   filter(height == growth) %>% 
   mutate_at(.vars = vars( height, growth, diameter, width, damage, 
                           vigour), ~replace(., nchar(.) >= 1, NA)) %>% 
   mutate(spp = ifelse(nchar(spp) >= 1, "0" , spp)) %>% 
   select(-germ_id) %>% 
   distinct() %>% 
  mutate(count = 0)
```
 
 #Now lets add this back in 
 
```{r}
zero_plots_final1 <- bind_rows(zero_plots_final, Zeroplots_germinants)
```

Lets make a new count for the DF
 
```{r}
 regen_data_ready_count1 <- regen_data %>% 
   filter(!plot %in% zero_plots_final1$plot) %>% 
   filter(spp == "Fd", regn_type == "N") %>% 
   group_by(gap, plot) %>% 
   summarise(count = n()) 
```
 
Now we create a new data frame to merge the zero_plots

```{r}
 zero_counts1 <- zero_plots_final1 %>% 
    select(gap, plot) %>% 
    mutate(count = 0)
```
 
Now I will bind the 2 data frames 

```{r}
regen_by_plot1 <- bind_rows(regen_data_ready_count1, zero_counts1) %>% 
    distinct() 
```
 
Now I will convert to total per hectare
```{r}
regen_by_plot1 <- regen_by_plot1 %>% 
    mutate(tph = count * 1000)
View(regen_by_plot1)
```

Now I will find density per gap
```{r}
regen_by_gap1 <- regen_by_plot1 %>% 
    group_by(gap) %>% 
    summarise(tphg = sum(tph), mean_density = mean(tph), sem_density = sd(tph)/sqrt(n()))
View(regen_by_gap1)
```
 
Now that it seems we have the right data, lets proceed to plot the scatter plots.
Since we already have the data on the LAI 4&5 for both S and N plots, we will proceed to join them to the new data we have.
 
For salvage LAI and plots
```{r}

 S_regen_LAI1 <- regen_by_plot1 %>% left_join(regen_LAIS) %>% 
    filter (gap_type =="S")
```

For natural LAI and plots
```{r}
N_regen_LAI1 <-  regen_by_plot1 %>% left_join(regen_LAIN) %>% 
    filter(gap_type =="N")
```
 
Now we plot scatter plots for both LAI Salavage
```{r scatter plot tph vrs lai S}
SScplot_LAI4R1 <- plot(tph ~ lai_4_ring, data = S_regen_LAI1)
SScplot_LAI5R1 <- plot (tph ~ lai_5_ring, data = S_regen_LAI1)

ggplot(S_regen_LAI1, aes(lai_4_ring, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salvage", x ="LAI", y= "Total trees per hectare")

ggplot(S_regen_LAI1, aes(lai_5_ring, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salvage", x ="LAI", y= "Total trees per hectare")

```

 
 
 #Now that we have the salavage LAI's lets plot that of the natural.
```{r scater plot tph vrs lai S}
 NScplot_4LAI1 <- plot(tph ~ lai_4_ring, data= N_regen_LAI1)
 NScplot_5LAI1 <- plot(tph ~lai_5_ring, data= N_regen_LAI1)
 
 ggplot(N_regen_LAI1, aes(lai_4_ring, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="LAI", y= "Total trees per hectare")

ggplot(N_regen_LAI1, aes(lai_5_ring, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="LAI", y= "Total trees per hectare")
```


#We have the scatter plots now. Before we continue let load the VRP data and find the summary on the species. First we clean names
```{r}
clean_names <- function(.data, unique = FALSE) {
    n <- if (is.data.frame(.data)) colnames(.data) else .data
    
    n <- gsub("%+", "_pct_", n)
    n <- gsub("\\$+", "_dollars_", n)
    n <- gsub("\\++", "_plus_", n)
    n <- gsub("-+", "_minus_", n)
    n <- gsub("\\*+", "_star_", n)
    n <- gsub("#+", "_cnt_", n)
    n <- gsub("&+", "_and_", n)
    n <- gsub("@+", "_at_", n)
    
    n <- gsub("[^a-zA-Z0-9_]+", "_", n)
    n <- gsub("([A-Z][a-z])", "_\\1", n)
    n <- tolower(trimws(n))
    
    n <- gsub("(^_+|_+$)", "", n)
    
    n <- gsub("_+", "_", n)
    
    if (unique) n <- make.unique(n, sep = "_")
    
    if (is.data.frame(.data)) {
       colnames(.data) <- n
       .data
    } else {
       n
    }
 }

```

Then we load the data VRP data
```{r}
regen_data2 <- read.csv('regen_data2.csv') %>% 
    clean_names() %>% 
    separate(gap_no, c("gap","plot")) 
 colnames(regen_data2)
```

To obtain the sum and mean of species  for both gap types we have: 
```{r Salvage Vigour per plot}
S_vrp_summarySplots <- regen_data2 %>%
  group_by(gap, plot, spp, n) %>% 
  filter(gap_type =='S')%>% 
  mutate(baf = n*3) %>% 
  group_by(gap, plot) %>% 
  mutate(total = sum(baf)) %>% 
  mutate(percent = (baf/total) * 100)
  view(S_vrp_summarySplots) 
```

For the Salvage gap
```{r Salvage vigour per gap}
S_vrp_summarySgap <- S_vrp_summarySplots %>% 
  group_by(gap, spp) %>% 
  mutate(baf = n*3) %>% 
  filter(baf != "NA") %>%
  summarise(bafn = sum(baf), sumn =sum(n), mean_baf = mean(baf),sem_baf =sd(baf)/sqrt(n()))%>%
  mutate(total_meanbaf = sum(mean_baf)) %>% 
  mutate(percent= (mean_baf/total_meanbaf)*100)
view(S_vrp_summarySgap)
```

For Salavge VRP general characteristics
```{r Salavge VRP general characteristics}
S_vrp_summarySStats <- S_vrp_summarySplots %>% 
  group_by(spp) %>% 
  mutate(baf = n*3) %>% 
  filter(baf != "NA") %>%
  summarise(bafn = sum(baf), sumn =sum(n), mean_baf = mean(baf),sem_baf =sd(baf)/sqrt(n()))%>%
  mutate(total_meanbaf = sum(mean_baf)) %>% 
  mutate(percent= (mean_baf/total_meanbaf)*100) %>% 
  filter(mean_baf!= "0")
view(S_vrp_summarySStats)
```


Now we do the same for the natural plot VRP

```{r natural plot VRP}
N_vrp_summaryplots <- regen_data2 %>% 
  group_by(gap, plot, spp, vrp_baf3, n) %>% 
  filter(gap_type == 'N') %>% 
  mutate(baf = n*3) %>% 
  group_by(gap, plot) %>% 
  mutate(total = sum(baf)) %>% 
  mutate(percent = (baf/total) * 100)
view(N_vrp_summaryplots)
```

 For the natural gap VRP 
```{r natural gap VRP }
N_vrp_summarygap <- N_vrp_summaryplots %>% 
 group_by(gap, spp) %>% 
  mutate(baf = n*3) %>% 
  filter(baf != "NA") %>%
  summarise(bafn = sum(baf), sumn =sum(n), mean_baf = mean(baf),sem_baf =sd(baf)/sqrt(n()))%>%
  mutate(total_meanbaf = sum(mean_baf)) %>% 
  mutate(percent= (mean_baf/total_meanbaf)*100)
view(N_vrp_summarygap)
```

Natural VRP general characteristics 
```{r}
N_vrp_summaryNStats <- N_vrp_summaryplots %>% 
 group_by(spp) %>% 
  mutate(baf = n*3) %>% 
  filter(baf != "NA") %>%
  summarise(bafn = sum(baf), sumn =sum(n), mean_baf = mean(baf),sem_baf =sd(baf)/sqrt(n()))%>%
  mutate(total_meanbaf = sum(mean_baf)) %>% 
  mutate(percent= (mean_baf/total_meanbaf)*100) %>% 
  filter(mean_baf!= "0")
view(N_vrp_summaryNStats)
 
```
Since I couldnt plot bar graphs for the VRP data. Lets continue. 
Lets work on the other light indices  and other growth parameters (height and height growth). 

```{r}
regen_data_ready_count2 <- regen_data %>% 
   filter(!plot %in% zero_plots_final1$plot) %>% 
   filter(spp == "Fd", regn_type == "N") %>% 
   group_by(gap, plot)
  
```

We need the zeros in the data 
```{r}
zero_counts2 <- zero_plots_final1 %>% 
    mutate(count = 0)
```

Merging the 2 dataframes to obtain data on the growth parameters.
 
```{r}
regen_by_plot2 <- bind_rows(regen_data_ready_count2, zero_counts2) %>% 
    distinct()
```

Now that we have the growth parameters and the zeros lets work on the gap types first.

Salvage regn by plot 2

```{r Salvage regn by plot 2}
regen_by_plot2S <- regen_by_plot2 %>% 
  filter(gap_type =="S") %>% 
  group_by(gap, plot) %>% 
  distinct()  %>% 
  mutate(growth = as.numeric(growth)) 
```

Natural regn by plot 

```{r Natural regn by plot 2}
regen_by_plot2N <- regen_by_plot2 %>% 
  filter(gap_type=="N") %>% 
  distinct() %>% 
  mutate(growth = as.numeric(growth))
view(regen_by_plot2N)
```

First lets deal with the mortaltity for Salvage per gap

```{r mortaltity for Salvage per gap}
mortS <- regen_by_plot2S%>% 
select(gap, plot, spp, status) %>% 
  filter(spp=="Fd") %>% 
  mutate(new_status= ifelse(status=="0", 0, status * 1))

mortzero <- mortS %>% 
  filter(spp==0) %>% 
  mutate(new_status = ifelse(spp == "0", NA, status * NA)) 

mortSnew <- bind_rows(mortS, mortzero)%>%
  distinct()  

morrtSnewG <- mortSnew %>% 
group_by(gap,new_status) %>%
summarise(count= n()) %>% 
  filter(new_status=="0") %>% 
  mutate(tph = count * 1000)
View(morrtSnewG)

MortSgeneral <- mortSnew %>%
  group_by(new_status) %>%
summarise(count= n()) %>%
  mutate(total = sum(count)) %>%  
  mutate(Percent=(count/total)* 100)
view(MortSgeneral)
```

Now mortatlity for Natural per gap

```{r mortatlity for Natural per gap}
mortN <- regen_by_plot2N%>% 
select(gap, plot, spp, status) %>% 
  filter(spp=="Fd") %>% 
  mutate(new_status= ifelse(status=="0", 0, status * 1))

mortzeroN <- mortN %>% 
  filter(spp==0) %>% 
  mutate(new_status = ifelse(spp == "0", NA, status * NA)) 

mortNnew <- bind_rows(mortN, mortzeroN)%>%
  distinct()

mortNnewG <- mortNnew %>% 
group_by(gap,new_status) %>%
summarise(count= n()) %>% 
  filter(new_status=="0") %>% 
  mutate(tph = count * 1000)
View(mortNnewG)

MortNgeneral <- mortNnew %>%
  group_by(new_status) %>%
summarise(count= n()) %>%
  mutate(total = sum(count)) %>%  
  mutate(Percent=(count/total)* 100)
view(MortNgeneral)
```

 Now that we have the mortality, lets move on to the height, growth, dameter and crown

For Height Salavage
```{r Height Salvage}
HsummaryS <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, height) %>%
  group_by(gap,plot) %>% 
  summarise(count=n(), meanH = mean(height, na.rm = TRUE), semH =sd(height, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() %>% 
  filter(meanH != 0)
```

For Height Summary Natural
```{r Height Summary Natural}
HsummaryN <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, height) %>%
  group_by(gap,plot) %>% 
  summarise(count=n(), meanH = mean(height, na.rm = TRUE), semH =sd(height, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() %>% 
  filter(meanH != 0)
```

Now for the height growth Salavage..... Check mean (growth)
```{r height growth Salavage}
GsummaryS <- regen_by_plot2S %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, growth) %>%
  group_by(gap,plot) %>% 
  mutate(growth = as.numeric(growth)) %>%
  summarise(count=n(), meanG = mean(growth, na.rm = TRUE), semG =sd(growth, na.rm = TRUE)/sqrt(n())) %>% 
  distinct()%>% 
  filter(meanG != 0)
```

For Height Growth Natural.... check mean(growth)
```{r Height Growth Natural}
GsummaryN <- regen_by_plot2N %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, growth) %>%
  group_by(gap, plot) %>% 
  mutate(growth = as.numeric(growth)) %>%
  summarise(count=n(), meanG = mean(growth, na.rm = TRUE), semG =sd(growth, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() %>% 
  filter(meanG != 0)
```

For Diameter Salavage 
```{r Diameter Salavage}
DsummaryS <- regen_by_plot2S %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, diameter) %>%
  group_by(gap, plot) %>%
  distinct() %>% 
  summarise(count=n(), meanD = mean(diameter, na.rm = TRUE), semD =sd(diameter, na.rm = TRUE)/sqrt(n())) %>% 
  filter(meanD !="NA") %>% 
  filter(meanD != 0)
```

For Diameter Natural
```{r  Diameter Natural}
DsummaryN <- regen_by_plot2N %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, diameter) %>%
  group_by(gap, plot) %>%
  distinct() %>% 
  summarise(count=n(), meanD = mean(diameter, na.rm = TRUE), semD =sd(diameter, na.rm = TRUE)/sqrt(n())) %>%  
   filter(meanD !="NA") %>% 
  filter(meanD != 0)
```

For Crown width Salavge... Check mean width
```{r Crown width Salavge}
WsummaryS <- regen_by_plot2S %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, width) %>%
  group_by(gap,plot) %>% 
  mutate(width = as.numeric(width)) %>% 
  summarise(count=n(), meanW = mean(width, na.rm = TRUE), semW =sd(width, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() %>% 
  filter(meanW != 0)
```


For Crown Width Natural .... Check mean width

```{r Crown Width Natural}
WsummaryN <- regen_by_plot2N %>% 
  filter(regn_type =="N") %>% 
  select(gap, plot, width) %>%
  group_by(gap,plot) %>%
  mutate(width = as.numeric(width)) %>%
  summarise(count=n(), meanW = mean(width, na.rm = TRUE), semW =sd(width, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() %>% 
  filter(meanW != 0)
```

Now lets deal with the Damage and Vigour 

For Damage Salvage
```{r  Damage Salvage per plot}
DAsummarySgap <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, damage) %>%
  group_by(gap,damage) %>% 
  distinct() %>%
  filter(damage!="NA") %>% # this part was to remove all plots with NA values*** 
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummarySgap)
```

Damage Salvage per plot
```{r}
DAsummarySplot <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, damage) %>%
  group_by(gap,plot,damage) %>% 
  distinct() %>%
  filter(damage!="NA") %>% # this part was to remove all plots with NA values*** 
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummarySplot)
```

Damage Salvage general statistics 
```{r Damage Salvage general statistics }
DAsummarySStats <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(damage) %>%
  group_by(damage) %>% 
  distinct() %>%
  filter(damage!="NA") %>% # this part was to remove all plots with NA values***
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummarySStats)
```

Now for Damage Natural

```{r Damage Natural per plot}
DAsummaryNplots <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, damage) %>%
  group_by(gap,plot,damage) %>% 
  distinct() %>%
   filter(damage!="NA") %>% # this part was to remove all plots with NA values*** 
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummaryNplots)
```

Damage natural per gap
```{r Damage natural per gap}
DAsummaryNgap <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, damage) %>%
  group_by(gap, damage) %>% 
  distinct() %>%
   filter(damage!="NA") %>% # this part was to remove all plots with NA values*** 
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummaryNgap)
```

Damage natural general statistics 
```{r Damage natural general statistics}
DAsummaryNStats <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(damage) %>%
  group_by(damage) %>% 
  distinct() %>%
  filter(damage!="NA") %>% # this part was to remove all plots with NA values***
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(DAsummaryNStats)
```

Now that we have the damage sorted out, let's work on the vigour

For the vigour Salvage

```{r vigour Salvage per plot}
VsummarySplots <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, vigour) %>%
  group_by(gap, plot, vigour) %>% 
  summarise(count=n()) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% #this part was to remove all plots with NA values***
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_vigour = (count/Count_total) * 100)
view(VsummarySplots)
```

Vigour Salvage per gap 
```{r Vigour Salvage per gap }
VsummarySgap <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, vigour) %>%
  group_by(gap,vigour) %>% 
  summarise(count=n()) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% # this part was to remove all plots with NA values***
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_vigour = (count/Count_total) * 100)
view(VsummarySgap)
```

Vigour slavage general summary 
```{r Vigour slavage general summary }
VsummarySStats <- regen_by_plot2S %>%
  filter(regn_type =="N") %>% 
  select(vigour) %>%
  group_by(vigour) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% # this part was to remove all plots with NA values***
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(VsummarySStats)
```

For the vigour Natural per plot 

```{r vigour Natural per plot }
VsummaryNplots <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, vigour) %>%
  group_by(gap, plot,vigour) %>% 
  summarise(count=n()) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% # this part was to remove all plots with NA values***
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_vigour = (count/Count_total) * 100)
view(VsummaryNplots)
```

Vigour Natural per gap
```{r Vigour Natural per gap}
VsummaryNgap <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, vigour) %>%
  group_by(gap,vigour) %>% 
  summarise(count=n()) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% # this part was to remove all plots with NA values***
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_vigour = (count/Count_total) * 100)
view(VsummaryNgap)
```

Vigour Natural general characteristics 
```{r VsummaryNStats}
VsummaryNStats <- regen_by_plot2N %>%
  filter(regn_type =="N") %>% 
  select(vigour) %>%
  group_by(vigour) %>% 
  distinct() %>%
  filter(vigour!="NA") %>% # this part was to remove all plots with NA values***
  summarise(count=n()) %>%
  mutate(Count_total = sum(count)) %>%
  mutate(Percent_damage = (count/Count_total) * 100)
view(VsummaryNStats)
```

Now lets work on the GLA openness and above and below ground irradiances and hope to find some relationship

lets begin with Openness for Salavge

```{r OpennnessS}
OpennnessS <- regen_data1N %>%
  group_by(gap, plot) %>% 
  filter(gap_type=="S") %>% 
  select(x_site_open) 
  
```

Openness for Natural
```{r OpennnessN}
OpennnessN <- regen_data1N %>%
  group_by(gap, plot) %>% 
  filter(gap_type=="N") %>% 
  select(x_site_open) 
```

Now lets deal with the above-and-below irradiances
in this instatnce we we will conside the trans total and the % trans total

For the irradiance Salavge 

```{r IrradianceS}
IrradianceS <- regen_data1N %>%
  group_by(gap, plot) %>% 
  filter(gap_type=="S") %>% 
  select("trans_tot", "x_trans_tot") 
```

For irradiance natural

```{r IrradianceN}
IrradianceN <- regen_data1N %>%
  group_by(gap, plot) %>% 
  filter(gap_type=="N") %>% 
  select("trans_tot", "x_trans_tot")
```

Now we left join the different informations to obtain our scatter plots.

Height

```{r}
H_open_S <- HsummaryS %>% left_join(OpennnessS)
H_open_N <- HsummaryN %>% left_join(OpennnessN)
H_irra_S <- HsummaryS %>% left_join(IrradianceS)
H_irra_N <- HsummaryN %>% left_join(IrradianceN)
H_LAI_S <- HsummaryS %>% left_join(regen_LAIS)
H_LAI_N <- HsummaryN %>% left_join(regen_LAIN)
```

Growth

```{r}
G_open_S <- GsummaryS %>% left_join(OpennnessS)
G_open_N <- GsummaryN %>% left_join(OpennnessN)
G_irra_S <- GsummaryS %>% left_join(IrradianceS)
G_irra_N <- GsummaryN %>% left_join(IrradianceN)
G_LAI_S <- GsummaryS %>% left_join(regen_LAIS)
G_LAI_N <- GsummaryN %>% left_join(regen_LAIN)
```

Diameter

```{r}
D_open_S <- DsummaryS %>% left_join(OpennnessS)
D_open_N <- DsummaryN %>% left_join(OpennnessN)
D_irra_S <- DsummaryS %>% left_join(IrradianceS)
D_irra_N <- DsummaryN %>% left_join(IrradianceN)
D_LAI_S <- DsummaryS %>% left_join(regen_LAIS)
D_LAI_N <- DsummaryN %>% left_join(regen_LAIN)
```

Width

```{r}
W_open_S <- WsummaryS %>% left_join(OpennnessS)
W_open_N <- WsummaryN %>% left_join(OpennnessN)
W_irra_S <- WsummaryS %>% left_join(IrradianceS)
W_irra_N <- WsummaryN %>% left_join(IrradianceN)
W_LAI_S <- WsummaryS %>% left_join(regen_LAIS)
W_LAI_N <- WsummaryN %>% left_join(regen_LAIN)
```
 
 Now lets try and plot some diagrams Using OPENNESS
 
Height vrs Openness

```{r} 
ggplot(H_open_S, aes(x_site_open, meanH)) +
  geom_point() +
  theme_classic()+
   labs(title= "Salvage", x= "%Site openess", y = "Total height (cm)")
```

```{r Natural}
ggplot(H_open_N, aes(x_site_open, meanH)) +
  geom_point() +
  theme_classic()+
  labs(title= "Natural", x= "%Site openess", y = "Total height (cm)")
```

Growth vrs Openness

```{r Salvage_1}
ggplot(G_open_S, aes(x_site_open, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "%Site openess", y = "Height increment (cm)")
```

```{r Natural_1}
ggplot(G_open_N, aes(x_site_open, meanG)) +
geom_point() +
theme_classic()+
   labs(title= "Natural", x= "%Site openess", y = "Height increment (cm)")
```

Diameter vrs Openness

```{r Salvage_2}
ggplot(D_open_S, aes(x_site_open, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "%Site openess", y = "Diameter (cm)")
```

```{r Natural_2}
ggplot(D_open_N, aes(x_site_open, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "%Site openess", y = "Diameter (cm)")
```

Crown width vrs Openness

```{r}
ggplot(W_open_S, aes(x_site_open, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "%Site openess", y = "Crown width (cm)")
```

```{r}
ggplot(W_open_N, aes(x_site_open, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Natural", x= "%Site openess", y = "Crown width (cm)")

```

Now lets try and plot some diagrams Using IRRADIANCE (trans_tot & X_trans_tot)

Height vrs Irradiance  

```{r}
ggplot(H_irra_S, aes(trans_tot, meanH)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "Transmitted total radiation", y = "Total height (cm)")

ggplot(H_irra_S, aes(trans_tot, meanH)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "%Transmitted total radiation", y = "Total height (cm)")
```


```{r }
ggplot(H_irra_N, aes(trans_tot, meanH)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "Transmitted total radiation", y = "Total height (cm)")

ggplot(H_irra_N, aes(x_trans_tot, meanH)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "%Transmitted total radiation", y = "Total height (cm)")
```

Growth vrs Irradiance 

```{r Growth vrs Irradiance }
ggplot(G_irra_S, aes(trans_tot, meanG)) +
geom_point() +
theme_classic()+
labs(title= "Salvage", x= "Transmitted total radiation", y = "Height increment (cm)")

ggplot(G_irra_S, aes(x_trans_tot, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "%Transmitted total radiation", y = "Height increment(cm)")
```

```{r }
ggplot(G_irra_N, aes(trans_tot, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "Transmitted total radiation", y = "Height increment(cm)")

ggplot(G_irra_N, aes(x_trans_tot, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "%Transmitted total radiation", y = "Height increment(cm)")
```

Diameter vrs Irradiance 
Salvage
```{r Diameter vrs Irradiance}
ggplot(D_irra_S, aes(trans_tot, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "Transmitted total radiation", y = "Diameter (cm)")

ggplot(D_irra_S, aes(x_trans_tot, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "%Transmitted total radiation", y = "Diameter (cm)")
```

Natural
```{r}
ggplot(D_irra_N, aes(trans_tot, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "Transmitted total radiation", y = "Diameter (cm)")

ggplot(D_irra_N, aes(x_trans_tot, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "%Transmitted total radiation", y = "Diameter (cm)")
```

Crown width vrs Irradiance 
Salvage
```{r Crown width vrs Irradiance }
ggplot(W_irra_S, aes(trans_tot, meanW)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "Transmitted total radiation", y = "Crown width (cm)")

ggplot(W_irra_S, aes(x_trans_tot, meanW)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "%Transmitted total radiation", y = "Crown width (cm)")
```

Natural
```{r}
ggplot(W_irra_N, aes(trans_tot, meanW)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "Transmitted total radiation", y = "Crown width (cm)")

ggplot(W_irra_N, aes(x_trans_tot, meanW)) +
geom_point() +
theme_classic()+
    labs(title= "Natural", x= "%Transmitted total radiation", y = "Crown width (cm)")
```

Height vrs LAI
Salavge
```{r Height vrs LAI}
ggplot(H_LAI_S, aes(lai_4_ring, meanH)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "LAI", y = "Total height (cm)")

ggplot(H_LAI_S, aes(lai_5_ring, meanH)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "LAI 5Rings", y = "Total height (cm)")
```

Natural
```{r }
ggplot(H_LAI_N, aes(lai_4_ring, meanH)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI", y = "Total height (cm)")
  

ggplot(H_LAI_N, aes(lai_5_ring, meanH)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI 5Rings", y = "Total height (cm)")
```

Growth vrs LAI
Salvage
```{r Growth vrs LAI }
ggplot(G_LAI_S, aes(lai_4_ring, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "LAI 4Rings", y = "Height increment (cm)")

ggplot(G_LAI_S, aes(lai_5_ring, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "LAI 5Rings", y = "Height increment (cm)")
```

Natural
```{r }
ggplot(G_LAI_N, aes(lai_4_ring, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI 4Rings", y = "Height increment (cm)")

ggplot(G_LAI_N, aes(lai_5_ring, meanG)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI 5Rings", y = "Height increment (cm)")
```

Diameter vrs LAI
Slavage
```{r Diameter vrs LAI}
ggplot(D_LAI_S, aes(lai_4_ring, meanD)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "LAI 4Rings", y = "Diameter (cm)")
   

ggplot(D_LAI_S, aes(lai_5_ring, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Salvage", x= "LAI 5Rings", y = "Diameter (cm)")
```

Natural
```{r diameter vrs LAI }
ggplot(D_LAI_N, aes(lai_4_ring, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI 4Rings", y = "Diameter (cm)")

ggplot(D_LAI_N, aes(lai_5_ring, meanD)) +
geom_point() +
theme_classic()+
  labs(title= "Natural", x= "LAI 5Rings", y = "Diameter (cm)")
```

Width vrs LAI
Salavage
```{r Width vrs LAI}
ggplot(W_LAI_S, aes(lai_4_ring, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "LAI 4Rings", y = "Crown width (cm)")

ggplot(W_LAI_S, aes(lai_5_ring, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Salvage", x= "LAI 5Rings", y = "Crown width (cm)")
```

Natural
```{r }
ggplot(W_LAI_N, aes(lai_4_ring, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Natural", x= "LAI 4Rings", y = "Crown width (cm)")

ggplot(W_LAI_N, aes(lai_5_ring, meanW)) +
geom_point() +
theme_classic()+
   labs(title= "Natural", x= "LAI 5Rings", y = "Crown width (cm)")
```

Let's work on the undeground cover, slope and aspect now. 

Underground cover

```{r Underground cover Salvage}
Undcover_summaryS <- regen_data1N %>% 
  select(gap, plot, gap_type, ms,ts,wd,gs,sb,fb,sl,rk,df) %>% 
  group_by(gap, plot, gap_type) %>% 
  filter(gap_type =="S") %>% 
   mutate(ms =as.numeric(ms), ts = as.numeric(ts), wd = as.numeric(wd), gs = as.numeric(gs), sb = as.numeric(sb), fb = as.numeric(fb), sl = as.numeric(sl), rk = as.numeric(rk), df = as.numeric(df)) 
View(Undcover_summaryS)
```


```{r Underground cover Salvage}
Undcover_summaryN <- regen_data1N %>% 
  select(gap, plot, gap_type, ms,ts,wd,gs,sb,fb,sl,rk,df) %>% 
  group_by(gap, plot, gap_type) %>% 
  filter(gap_type =="N") %>% 
   mutate(ms =as.numeric(ms), ts = as.numeric(ts), wd = as.numeric(wd), gs = as.numeric(gs), sb = as.numeric(sb), fb = as.numeric(fb), sl = as.numeric(sl), rk = as.numeric(rk), df = as.numeric(df))
view(Undcover_summaryN)
```

Regen vrs underground cover

```{r Regen vrs underground cover}
Regen_by_coverS <- regen_by_plot1 %>%left_join(Undcover_summaryS) %>% 
  filter(gap_type !="NA") 

  
Regen_by_coverN <- regen_by_plot1 %>%left_join(Undcover_summaryN) %>% 
  filter(gap_type !="NA")
```

Now let's work on the Aspect and Slope 
Salvage

```{r Salvage Aslo}
Aslo_summaryS <- regen_data1N %>% 
  select(gap, plot, gap_type, aspect, slope) %>% 
  group_by(gap, plot) %>%
  filter(gap_type =="S") 
 
```

Transforming the aspect Salvage
```{r Transforming the aspect Salvage}
New_aspectS <- Aslo_summaryS %>% 
  select(gap_type, aspect) %>% 
  mutate(aspectT = cos(rad(0-aspect))+1)
```

Natural
```{r Natural Aslo}
Aslo_summaryN <- regen_data1N %>% 
  select(gap, plot, gap_type, aspect, slope) %>% 
  group_by(gap, plot) %>%
  filter(gap_type =="N")
```
Transforming aspect Natural
```{r}
New_aspectN <- Aslo_summaryN %>% 
  select(gap_type, aspect) %>% 
  mutate(aspectT = cos(rad(0-aspect))+1)
```

Plotting aspect 
```{r Plot of aspects}
ggplot(New_aspectS, aes(aspect, aspectT)) +
geom_point() +
theme_classic()+
  labs(title = "Salvage", x = "Measured aspect", y = "Transformed aspect")

ggplot(New_aspectN, aes(aspect, aspectT)) +
geom_point() +
theme_classic()+
  labs(title = "Natural", x = "Measured aspect", y = "Transformed aspect")
```


Aspect and slope vrs Regen

```{r Aspect and slope vrs Regen}
Regen_by_Aslo_S <- regen_by_plot1 %>% left_join(Aslo_summaryS) %>% 
  filter(gap_type =="S")

Regen_by_Aslo_N <- regen_by_plot1 %>% left_join(Aslo_summaryN) %>% 
  filter(gap_type !="NA")

Regen_by_NewaspectS <- regen_by_plot1 %>% left_join(New_aspectS) %>% 
   filter(gap_type =="S")

Regen_by_NewaspectN <- regen_by_plot1 %>% left_join(New_aspectN) %>% 
   filter(gap_type !="NA")
``` 

Plot for aspect and slope relative to regeneration 
Salvage
```{r}
ggplot(Regen_by_NewaspectS, aes(aspectT, tph)) +
geom_point() +
theme_classic() +
labs(title = "Salvage", x = "Transformed aspect", y = "TPH")

ggplot(Regen_by_Aslo_S, aes(slope, tph)) +
geom_point() +
theme_classic()+
labs(title = "Salvage", x = "Slope", y = "TPH")
```

Natural 
```{r}
ggplot(Regen_by_NewaspectN, aes(aspectT, tph)) +
geom_point() +
theme_classic()+
labs(title = "Natural", x = "Transformed aspect", y = "TPH")

ggplot(Regen_by_Aslo_N, aes(slope, tph)) +
geom_point() +
theme_classic()+
labs(title = "Natural", x = "Slope", y = "TPH")
```

Could there be a possible relationship between height growth and slope and aspect?
Joining the datasets together

```{r}
Height_vrs_Aslo_S <- HsummaryS %>% left_join(Aslo_summaryS)
Height_vrs_Aslo_N <- HsummaryN %>% left_join(Aslo_summaryN)
Growth_vrs_Aslo_S <- GsummaryS %>% left_join(Aslo_summaryS)
Growth_vrs_Aslo_N <- GsummaryN %>% left_join(Aslo_summaryN)

Height_vrs_aspectS <- HsummaryS %>% left_join(New_aspectS)
Height_vrs_aspectN <- HsummaryN %>% left_join(New_aspectN)
Growth_vrs_aspectS <- GsummaryS %>% left_join(New_aspectS)
Growth_vrs_aspectN <- GsummaryN %>% left_join(New_aspectN)
```

Plots of height and growth vrs slope and aspect
Salvage
```{r }
ggplot(Height_vrs_aspectS, aes(aspectT, meanH, )) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Transformed aspect", y= "Total height (cm)")

ggplot(Height_vrs_Aslo_S, aes(slope, meanH)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Slope", y= "Total height (cm)")
```

Natural
```{r }
ggplot(Height_vrs_aspectN, aes(aspectT, meanH)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Transformed aspect", y= "Total height (cm)")

ggplot(Height_vrs_Aslo_N, aes(slope, meanH)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Slope", y= "Total height (cm)")
```

Now lets look at regen and underground cover plots and see if there is a relationship. Since I am unable to plot all the parameters at the same time. 
Lets look at them one after the other and see the resultant graphs they produce.

For Moss Slavage and natural

```{r}
ggplot(Regen_by_coverS, aes(ms, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Moss", y= "TPH")

ggplot(Regen_by_coverN, aes(ms, tph)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Moss", y= "TPH")
```

For Trees Slavage and natural

```{r}
ggplot(Regen_by_coverS, aes(as.numeric(ts), tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Trees", y= "TPH")

ggplot(Regen_by_coverN, aes(as.numeric(ts), tph)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Trees", y= "TPH")
```

For Wood Salvage and Natural
```{r}
ggplot(Regen_by_coverS, aes(wd, tph)) +
geom_point()+
theme_classic()+
   labs(title="Salvage", x= "Wood", y= "TPH")

ggplot(Regen_by_coverN, aes(wd, tph)) +
geom_point()+
theme_classic()+
    labs(title="Natural", x= "Wood", y= "TPH")
```

For Grass
```{r}
ggplot(Regen_by_coverS, aes(gs, tph)) +
geom_point()+
theme_classic()+
    labs(title="Salvage", x= "Grass", y= "TPH")

ggplot(Regen_by_coverN, aes(gs, tph)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Grass", y= "TPH")
```

For Shrub
```{r}
ggplot(Regen_by_coverS, aes(sb, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Shrub", y= "TPH")

ggplot(Regen_by_coverN, aes(sb, tph)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Grass", y= "TPH")
```

For Forbs
```{r}
ggplot(Regen_by_coverS, aes(fb, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Forbs", y= "TPH")

ggplot(Regen_by_coverN, aes(fb, tph)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Grass", y= "TPH")
```

For soil
```{r}
ggplot(Regen_by_coverS, aes(sl, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Soil", y= "TPH")

ggplot(Regen_by_coverN, aes(sl, tph)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Soil", y= "TPH")
```

For rock 
```{r}
ggplot(Regen_by_coverS, aes(rk, tph)) +
geom_point()+
theme_classic()+ 
  labs(title="Salvage", x= "Rock", y= "TPH")
  

ggplot(Regen_by_coverN, aes(rk, tph)) +
geom_point()+
theme_classic()+
  labs(title="Natural", x= "Rock", y= "TPH")
```

For Duff 
```{r}
ggplot(Regen_by_coverS, aes(df, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Duff", y= "TPH")

ggplot(Regen_by_coverN, aes(df, tph)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Duff", y= "TPH")
```


To determine the total height of seedling last year, I subtract the height increment from the total height to obtain last years height.

```{r}
LastyrHeightS <- regen_by_plot2S %>% 
  mutate(LstyrHgt = height - growth)
  view(LastyrHeightS)
  
LastyrHeightN <- regen_by_plot2N %>% 
  mutate(LstyrHgt = height - growth)
  view(LastyrHeightN)
```


```{r}
ggplot(LastyrHeightS, aes(LstyrHgt, growth))+
  geom_point()+
  theme_classic()+
  labs(title= "Salvage", x="Last year's total height (cm)", y= "Height increment (cm)")


ggplot(LastyrHeightN, aes(LstyrHgt, growth))+
  geom_point()+
  theme_classic()+
  labs(title= "Natural", x="Last year's total height (cm)", y= "Height increment (cm)")
```

Lets check for the summary statistics of the new height growth and see if it gives us a better plot

For Height Salavage
```{r}
HnewsummaryS <- LastyrHeightS %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, LstyrHgt) %>%
  group_by(gap,plot) %>% 
  summarise(count=n(), meanH = mean(LstyrHgt, na.rm = TRUE), semH =sd(LstyrHgt, na.rm = TRUE)/sqrt(n())) %>% 
  distinct()
```

For Height Summary Natural
```{r}
HnewsummaryN <- LastyrHeightN %>%
  filter(regn_type =="N") %>% 
  select(gap, plot, LstyrHgt) %>%
  group_by(gap,plot) %>% 
  summarise(count=n(), meanH = mean(LstyrHgt, na.rm = TRUE), semH =sd(LstyrHgt, na.rm = TRUE)/sqrt(n())) %>% 
  distinct() 

```

Using means for last year height and increment
```{r Using means for last year height and increment}
Lst_vrs_HI_S <- HnewsummaryS %>%left_join(GsummaryS)
Lst_vrs_HI_N <- HnewsummaryN %>%left_join(GsummaryN)
```

Now lets plot last year height and increment based on means 
```{r last year height and increment }
ggplot(Lst_vrs_HI_S, aes(meanH, meanG))+
  geom_point()+
  theme_classic()+
  labs(title= "Salvage", x="Last year's total height (cm)", y= "Height increment (cm)")

ggplot(Lst_vrs_HI_N, aes(meanH, meanG))+
  geom_point()+
  theme_classic()+
  labs(title= "Natural", x="Last year's total height (cm)", y= "Height increment (cm)")
```


The above graphs look somewhat linear, so lets generate a linear model for both of them and choose the best fit ones 
```{r Model 1 Salvage}
Model1S <- lm(growth ~ LstyrHgt, data = LastyrHeightS)
plot(Model1S)
abline(Model1S, col="blue")
Model1S
summary(Model1S)
AIC(Model1S)
```

```{r Model1 Natural}

Model1N <- lm(growth ~ LstyrHgt, data = LastyrHeightN)
plot(Model1N)
abline(Model1N, col="blue")
Model1N
summary(Model1N)
AIC(Model1N)
```

Model 2
```{r Model2S}
Model2S <- lm(meanH ~ meanG, data = Lst_vrs_HI_S)
plot(Model2S)
abline(Model2S, col="blue")
Model2S
summary(Model2S)
AIC(Model2S)
```

Natural
```{r Model 2N}
Model2N <- lm(meanH ~ meanG, data = Lst_vrs_HI_N)
plot(Model2N)
abline(Model2N, col="blue")
Model2N
summary(Model2N)
AIC(Model2N)
```

Lets join growth to underground cover

```{r}
Growth_by_coverS <- GsummaryS %>%left_join(Undcover_summaryS) %>% 
  filter(gap_type !="NA") 

  
Growth_by_coverN <- GsummaryN %>%left_join(Undcover_summaryN) %>% 
  filter(gap_type !="NA")
```


```{r}
ggplot(Growth_by_coverS, aes(ms, meanG)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Moss", y= "Height increment (cm)")

ggplot(Growth_by_coverN, aes(ms, meanG)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Moss", y= "Height increment (cm)")
```

NOw l am going to look at the influence of distance to edge of the plot
```{r}
neardist <- read_csv("near_distance.csv") %>% 
   clean_names() %>% 
    separate(plot_id, c("gap","plot"))
 colnames(neardist)
 summary(neardist)
```

Lets join the distance to TPH and height, growth, diameter and crown for the salvage 
```{r distance to TPH}
 dist_vrs_TPH_S <- (regen_by_plot1) %>%  left_join(Aslo_summaryS) %>% left_join(neardist) %>% 
  filter(gap_type == "S") %>% 
  filter(tph != "NA")
view(dist_vrs_TPH_S) 

dist_vrs_TPH_N <- (regen_by_plot1) %>%  left_join(Aslo_summaryN) %>% left_join(neardist) %>% 
  filter(gap_type == "N") %>% 
  filter(tph != "NA")
view(dist_vrs_TPH_N)
```

Now lets join the distance to growth variables salvage 
```{r distance to growth variables salvage}
dist_vrs_growthvarS <- neardist %>% left_join(HsummaryS) %>% left_join(GsummaryS) %>% left_join(DsummaryS) %>% left_join(WsummaryS)  

  
dist_vrs_growthvarN <- neardist %>% left_join(HsummaryN) %>% left_join(GsummaryN) %>% left_join(DsummaryN) %>% left_join(WsummaryN)   
```

Now lets plot some graphs 
```{r Distance vrs TPH}
ggplot(dist_vrs_TPH_S, aes(near_distance, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to gap edge (m)", y= "Total trees per hectare")

ggplot(dist_vrs_TPH_N, aes(near_distance, tph)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to gap edge (m)", y= "Total trees per hectare")
```

NOw lets see the growth parameters
```{r distance vrs height}
ggplot(dist_vrs_growthvarS, aes(near_distance, meanH)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (m)", y= "Total height (cm)") +
  xlim(0, 6.5)
 

ggplot(dist_vrs_growthvarN, aes(near_distance, meanH)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (m)", y= "Total height (cm)") +
  xlim(0, 4)

```

```{r diatance vrs growth}
ggplot(dist_vrs_growthvarS, aes(near_distance, meanG)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (m)", y= "Height increment (cm)")  +
  xlim(0, 7.5)
 

ggplot(dist_vrs_growthvarN, aes(near_distance, meanG)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (m)", y= "Height increment (cm)")  +
  xlim(0, 5)
 
```

```{r distance vrs diameter}
ggplot(dist_vrs_growthvarS, aes(near_distance, meanD)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (m)", y= "Diameter (cm)")  +
  xlim(0, 7)
  

ggplot(dist_vrs_growthvarN, aes(near_distance, meanD)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (m)", y= "Diameter (cm)")  +
  xlim(0, 4)
  
```

```{r distance vrs width}
ggplot(dist_vrs_growthvarS, aes(near_distance, meanW)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (m)", y= "Crown width (cm)")  +
  xlim(0, 7)
  

ggplot(dist_vrs_growthvarN, aes(near_distance, meanW)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (m)", y= "Crown width (cm)")  +
  xlim(0, 4)
 
```

Lets find the relation between LAI and the distance to edge 
```{r Joining LAI and the distance to edge}
Dist_vrs_LAI <- neardist %>%  left_join(regen_data1N) 
view(Dist_vrs_LAI)

Dist_vrs_LAI_S <- Dist_vrs_LAI %>% 
  filter(gap_type == "S")
view(Dist_vrs_LAI_S)

Dist_vrs_LAI_N <- Dist_vrs_LAI %>% 
  filter(gap_type == "N")
view(Dist_vrs_LAI_N)
```
 Lets find scatter plots 
 
```{r Distance to edge vrs LAI4}
ggplot(Dist_vrs_LAI_S, aes(near_distance, lai_4_ring)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (cm)", y= "LAI")
  

ggplot(Dist_vrs_LAI_N, aes(near_distance, lai_4_ring)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (cm)", y= "LAI")
```
 
```{r distance vrs LAI 5R}
ggplot(Dist_vrs_LAI_S, aes(near_distance, lai_5_ring)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (cm)", y= "LAI 5rings")
  

ggplot(Dist_vrs_LAI_N, aes(near_distance, lai_5_ring)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (cm)", y= "LAI 5rings")
```

Now lets find scatter plot of height vrs crown 
```{r Height vrs crown relationship}
Hgt_vrs_crownS <- HsummaryS %>% left_join(WsummaryS)
view(Hgt_vrs_crownS)

Hgt_vrs_crownN <- HsummaryN %>% left_join(WsummaryN)
view(Hgt_vrs_crownN)
```
 Now lets plot height vrs crown and see if there is any rekationship
 
```{r}
ggplot(Hgt_vrs_crownS, aes(meanW, meanH)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Crown width (cm)", y= "Total height (cm)")
  

ggplot(Hgt_vrs_crownN, aes(meanW, meanH)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Crown width (cm)", y= "Total height (cm)")
```

Now lets see if there is a relationship between BAF and distance to edge 
```{r}
BAF_vrs_distS <- S_vrp_summarySplots %>% left_join(neardist) 
view(BAF_vrs_distS)

BAF_vrs_distN <- N_vrp_summaryplots %>% left_join(neardist) 
view(BAF_vrs_distN)
```

Now lets plot and see if there is a relationship
```{r BA vrs distance to edge}
ggplot(BAF_vrs_distS, aes(near_distance, total)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Distance to edge (m)", y= "Basal area (m^2)")
  

ggplot(BAF_vrs_distN, aes(near_distance, total)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Distance to edge (m)", y= "Basal area (m^2)")

```

Now lets find out if there is a relationship between BA and growth variables
```{r Ba vrs growth }
BAF_vrs_growthvarS <- S_vrp_summarySplots %>% left_join(regen_by_plot1) %>% left_join(HsummaryS) %>% left_join(GsummaryS) %>% left_join(DsummaryS) %>% left_join(WsummaryS)
view(BAF_vrs_growthvarS)

BAF_vrs_growthvarN <- N_vrp_summaryplots %>% left_join(regen_by_plot1) %>% left_join(HsummaryN) %>% left_join(GsummaryN) %>% left_join(DsummaryN) %>% left_join(WsummaryN)
view(BAF_vrs_growthvarN)
```

Now lets plot and see if there is any relationship between BA and growth variables 
```{r BA vrs TPH}
ggplot(BAF_vrs_growthvarS, aes(total, tph)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y= "Total trees per hectare")
  

ggplot(BAF_vrs_growthvarN, aes(total, tph)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Basal area (m^2)", y= "Total trees per hectare")
```

```{r BA vrs height}
ggplot(BAF_vrs_growthvarS, aes(total, meanH)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y= "Total height (cm)")
  

ggplot(BAF_vrs_growthvarN, aes(total, meanH)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Basal area (m^2)", y= "Total height (cm)")
```

```{r BA vrs Growth }
ggplot(BAF_vrs_growthvarS, aes(total, meanG)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y= "Height increment (cm)")
  

ggplot(BAF_vrs_growthvarN, aes(total, meanG)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Basal area (m^2)", y= "Height increment (cm)")
```

```{r BA vrs diameter}
ggplot(BAF_vrs_growthvarS, aes(total, meanD)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y= "Diameter (cm)") 

  

ggplot(BAF_vrs_growthvarN, aes(total, meanD)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Basal area (m^2)", y= "Diameter (cm)")
  
```

```{r BA vrs crown}
ggplot(BAF_vrs_growthvarS, aes(total, meanW)) +
geom_point()+
theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y= "Crown width (cm)") 

  

ggplot(BAF_vrs_growthvarN, aes(total, meanW)) +
geom_point()+
theme_classic()+
   labs(title="Natural", x= "Basal area (m^2)", y= "Crown width (cm)")
```

Now lets see if there is a signicant difference between TPH for the natural and salvaged plots 

```{r BOxplot of lai 4ring }
ggplot(regen_data1N, aes(as.factor(gap_type), lai_4_ring))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "LAI 4rings")

LAI_4rings_anova <- aov(lai_4_ring ~ gap_type, data = regen_data1N)
summary(LAI_4rings_anova)

```

```{r Boxplot of lai 5rings}
ggplot(regen_data1N, aes(as.factor(gap_type), lai_5_ring))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "LAI 5rings")

LAI_5rings_anova <- aov(lai_5_ring ~ gap_type, data = regen_data1N)
summary(LAI_4rings_anova)
```

 Now lets see tph
```{r tph boxplot  natural vrs salvage}
tph_vrs_regen <- regen_by_plot1 %>% left_join(regen_data1N) %>% 
select(tph, gap_type)
view(tph_vrs_regen)

ggplot(tph_vrs_regen, aes(as.factor(gap_type), tph))+
  geom_boxplot()+
  theme_classic()+
  labs(x="gap_type", y="TPH")

TPH_anova <- aov(tph ~ gap_type, data = tph_vrs_regen)
summary(TPH_anova)
```
 
Now lets see total radiation 
```{r  Total radiation boxplot  natural vrs salvage}
ggplot(regen_data1N, aes(as.factor(gap_type), trans_tot))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "Total transmitted radiation")

Totalradiation_anova <- aov(trans_tot ~ gap_type, data = regen_data1N)
summary(Totalradiation_anova)
```

Now lets see %openness 
```{r  %Openness boxplot  natural vrs salvage}
ggplot(regen_data1N, aes(as.factor(gap_type), x_site_open))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "%Opennness")

Openness_anova <- aov(x_site_open ~ gap_type, data = regen_data1N)
summary(Totalradiation_anova)
```

Now lets see %Total radiation 
```{r  %total radiation boxplot  natural vrs salvage}
ggplot(regen_data1N, aes(as.factor(gap_type), x_trans_tot))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "%Total transmitted radiation")

Openness_anova <- aov(x_trans_tot ~ gap_type, data = regen_data1N)
summary(Totalradiation_anova)
```

For growth variables 
```{r Growth variables boxplot Natural vrs Salvage}
GrowthVariables <- regen_by_plot2 %>%
  select(gap, plot, gap_type, height, growth, diameter, width) %>%
  group_by(gap,plot, gap_type) %>% 
  mutate(height = as.numeric(height), growth = as.numeric(growth), diameter = as.numeric(diameter), width = as.numeric(width), ) %>% 
  summarise(count=n(), meanH = mean(height, na.rm = TRUE), meanG = mean(growth, na.rm = TRUE), meanD = mean(diameter, na.rm = TRUE), meanW = mean(width, na.rm = TRUE)) %>% 
  distinct()
 
```

Now lets see height natural vrs salvage 
```{r height natural vrs salvage }
ggplot(GrowthVariables, aes(as.factor(gap_type), meanH))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "Total height (cm)")

Height_anova <- aov(meanH ~ gap_type, data = GrowthVariables)
summary(Height_anova)
```

```{r growth natural vrs salvage}
ggplot(GrowthVariables, aes(as.factor(gap_type), meanG))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "Height increment (cm)")

Growth_anova <- aov(meanG ~ gap_type, data = GrowthVariables)
summary(Growth_anova)
```

```{r diameter natural vrs salvage}
ggplot(GrowthVariables, aes(as.factor(gap_type), meanD))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "Diameter (cm)")

Diameter_anova <- aov(meanD ~ gap_type, data = GrowthVariables)
summary(Diameter_anova)
```

```{r crown natural vrs salvage}
ggplot(GrowthVariables, aes(as.factor(gap_type), meanW))+
  geom_boxplot()+
  theme_classic()+
  labs(x= "Gap type", y= "Crown width (cm)")

Width_anova <- aov(meanW ~ gap_type, data = GrowthVariables)
summary(Width_anova)
```

Test for the relationship between openness and LAI

```{r Relationship between openness and LAI}
LAI_vrs_Open <- regen_data1N %>% 
  select(x_site_open, lai_4_ring, lai_5_ring)

ggplot(LAI_vrs_Open, aes(x_site_open, lai_4_ring))+
   geom_point()+
  theme_classic()+
  labs(x= "%Openness", y= "LAI 4ring")

ggplot(LAI_vrs_Open, aes(x_site_open, lai_5_ring))+
  geom_point()+
  theme_classic()+
  labs(x= "%Openness", y= "LAI 5ring")
```

Lets see the relationship between distance to the edge_vrs_openness 
```{r distance to the edge_vrs_openness}
Dist_vrs_openness <- neardist %>% left_join(regen_data1N) %>% 
  select(near_distance, x_site_open, gap_type)

Dist_vrs_opennessS <- Dist_vrs_openness %>% 
  filter(gap_type == "S")

Dist_vrs_opennessN <- Dist_vrs_openness %>% 
  filter(gap_type == "N")
```
Now lets plot the graphs
```{r}
ggplot(Dist_vrs_opennessS, aes(near_distance, x_site_open))+
  geom_point() +
  theme_classic() +
  labs(title = "Salvage", x= "Distance to gap edge (cm)", y= "% Gap openness")

ggplot(Dist_vrs_opennessN, aes(near_distance, x_site_open))+
  geom_point() +
  theme_classic() +
  labs(title = "Natural", x= "Distance to gap edge (cm)", y= "% Gap openness")
```

Now lets transform the Azimuth 
```{r Transformed Azimuth}
Azimuth_trans <- neardist %>% 
  mutate(azimuth_T = cos(rad(0-azimuth1))+1)
view(Azimuth_trans)

Azimuth_vrs_growthS <- Azimuth_trans %>% left_join(HsummaryS) %>% left_join(GsummaryS) %>% left_join(DsummaryS) %>% left_join(WsummaryS) %>% 
  filter(count !="NA") 
view(Azimuth_vrs_growthS) 

Azimuth_vrs_growthN <- Azimuth_trans %>% left_join(HsummaryN) %>% left_join(GsummaryN) %>% left_join(DsummaryN) %>% left_join(WsummaryN) %>% 
  filter(count !="NA") 
view(Azimuth_vrs_growthN) 
```

Now lets plot relationships between transformed azimuth and growth var

```{r Azimuth vrs height}
ggplot(Azimuth_vrs_growthS, aes(azimuth_T, meanH)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salavge", x ="Azimuth", y= "Total height (cm)")

ggplot(Azimuth_vrs_growthN, aes(azimuth_T, meanH)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="Azimuth", y= "Total height (cm)")
```

```{r Azimuth vrs growth}
ggplot(Azimuth_vrs_growthS, aes(azimuth_T, meanG)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salavge", x ="Azimuth", y= "Heiught increment (cm)")

ggplot(Azimuth_vrs_growthN, aes(azimuth_T, meanG)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="Azimuth", y= "Height increment (cm)")
```

```{r Azimuth vrs diameter}
ggplot(Azimuth_vrs_growthS, aes(azimuth_T, meanD)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salavge", x ="Azimuth", y= "Diameter (cm)")

ggplot(Azimuth_vrs_growthN, aes(azimuth_T, meanD)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="Azimuth", y= "Diameter (cm)")
```

```{r Azimuth vrs width}
ggplot(Azimuth_vrs_growthS, aes(azimuth_T, meanW)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salavge", x ="Azimuth", y= "Crown width (cm)")

ggplot(Azimuth_vrs_growthN, aes(azimuth_T, meanW)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="Azimuth", y= "Crown width (cm)")
```

Transformed azimuth vrs TPH 
```{r}
AziTPH_S <- regen_by_plot1 %>% left_join(Azimuth_trans) %>% left_join(regen_data1N) %>% 
  filter(gap_type !="N")
view(AziTPH_S)

AziTPH_N <- regen_by_plot1 %>% left_join(Azimuth_trans) %>% left_join(regen_data1N) %>% 
  filter(gap_type !="S")
view(AziTPH_N)

ggplot(AziTPH_S, aes(azimuth_T, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Salavge", x ="Azimuth", y= "Total trees per hectare")

ggplot(AziTPH_N, aes(azimuth_T, tph)) +
  geom_point()+
  theme_classic()+
  labs(title ="Natural", x ="Azimuth", y= "Total trees per hectare")
```

Now that we have data on the dem lets see what we can find
```{r}
Demdata <- read.csv("dem.csv") %>% 
  clean_names() %>% 
    separate(plot_id, c("gap","plot"))
 colnames(Demdata)
 summary(Demdata)
```

Lets transform the aspect for dem1m and plot it and slope with growth variables
```{r}
Demdata <- Demdata %>% 
  mutate(aspect_dem11T = cos(rad(0-aspect_dem11))+1) %>% 
  mutate(aspect_dem21T = cos(rad(0-aspect_dem21))+1) %>% 
  mutate(aspect_dem51T = cos(rad(0-aspect_dem51))+1) %>% 
  mutate(aspect_dem101T = cos(rad(0-aspect_dem101))+1)
view(Demdata)
```

LAI vrs BA
```{r}
LAIvvrsBA_S <- regen_data1N %>% left_join(S_vrp_summarySplots) %>% 
filter(gap_type != "N")

LAIvvrsBA_N <- regen_data1N %>% left_join(N_vrp_summaryplots) %>% 
filter(gap_type != "S")
```

Now lets plot
```{r}
ggplot(LAIvvrsBA_N, aes(total, lai_4_ring)) +
geom_point()+
  theme_classic()+
  labs(title="Natural", x= "Basal area (m^2)", y="LAI")

ggplot(LAIvvrsBA_S, aes(total, lai_4_ring)) +
geom_point()+
  theme_classic()+
  labs(title="Salvage", x= "Basal area (m^2)", y="LAI")
```

