---
title: "Data Analysis_1"
author: "Emmanuel."
date: "07/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r}
library(tidyverse)
```


##Importing Data

```{r}
library(readxl)
Regen_data <- read_excel("C:/Users/adoasi94/Desktop/Regen_data_NewBackUp.xlsx", sheet = "REGEN")
View(Regen_data_NewBackUp)
head(Regen_data_NewBackUp)
```

```{r}
Regen_data_NewBackUp <- Regen_data_NewBackUp %>% 
  rename(plot = Gap_No) %>% 
  separate(plot, c("Gap", "Plot"))
```



```{r}
Salvage_regen_summary_plots <- Regen_data_NewBackUp %>% 
  filter(Gap_type == "S") %>% 
  group_by(Gap, Plot, Spp) %>% 
  summarise(count = n()) %>%
  mutate(TPH = ifelse(Spp == "0", 0, count * 1000)) 
view(Salvage_regen_summary_plots)
  

length(unique(Salvage_regen_summary_plots$Gap))
length(unique(Salvage_regen_summary_plots$Plot))
length(unique(Regen_data_NewBackUp$Plot))
length(unique(Regen_data_NewBackUp$Gap))
plot_test <- unique(Regen_data_NewBackUp)
View(Regen_data_NewBackUp)
plot_test <- unique(Regen_data_NewBackUp)
plot_test <- sort(unique(Regen_data_NewBackUp$Plot))
plot_test <- as.data.frame(as.numeric(sort(unique(Regen_data_NewBackUp$Plot))))
View(plot_test)
plot_test <- as.data.frame(plot =as.numeric(sort(unique(Regen_data_NewBackUp$Plot))))
plot_test <- as.data.frame(as.numeric(sort(unique(Regen_data_NewBackUp$Plot))))
View(plot_test)
plot_test <- plot_test %>% rename(plot = `as.numeric(sort(unique(Regen_data_NewBackUp$Plot)))`)
plot_test <- plot_test %>% arrange(plot)
plot_test$row_id <- rownames(plot_test$plot)
View(plot_test)
plot_test <- plot_test %>%  rownames_to_column("rowname")
plot_test <- plot_test %>% mutate(test = ifelse(plot == rowname, "Good", "Bad"))
```
```{r}
view(Salvage_regen_summary_plots)
SMeanSpecies <- Salvage_regen_summary_plots %>%
  group_by(Gap,Spp) %>% 
  summarise(mean_gap=mean(TPH))
View(SMeanSpecies)
```


Generating a table for the results
```{r}
Salvage_regen_summary_plots %>%
  group_by(Gap,Spp) %>% 
  summarise(mean_gap=mean(TPH)) %>% 
  mutate(mean_gap) %>% 
  
```

Determining number of species per hectare per plots 

```{r}
Natural_regen_summary_plots <- Regen_data_NewBackUp %>% 
  filter(Gap_type == "N") %>% 
  group_by(Gap, Plot, Spp) %>% 
  summarise(count = n()) %>%
  mutate(TPHN = ifelse(Spp == "0", 0, count * 1000)) 
view(Natural_regen_summary_plots)
```

Summary of Species per hectare per Natural gap 
```{r}
view(Natural_regen_summary_plots)
NMeanSpecies <- Natural_regen_summary_plots %>%
  group_by(Gap,Spp) %>% 
  summarise(mean_gap=mean(TPHN))
View(NMeanSpecies)
```

Total number of species per Salvage gap 

```{r}
view(Salvage_regen_summary_plots)
SMeanSpeciesTotal <- Salvage_regen_summary_plots %>%
  group_by(Gap) %>% 
  summarise(mean_gap1=mean(TPH))
```

Total number of species per Natural gap
```{r}
NMeanSpeciesTotal <- Natural_regen_summary_plots %>%
  group_by(Gap) %>% 
  summarise(mean_gap2=mean(TPHN))
```
Mean of salvage species per plot to match LAI 4R data

```{r}
view(Salvage_regen_summary_plots)
STotalSpeciesPlot <- Salvage_regen_summary_plots %>%
  group_by(Gap, Plot) %>% 
  summarise(mean_plot=mean(TPH)) 
```

Group LAI to match up Salvage regeneration data
Using PASG
```{r}
Regen_data_NewBackUp <- Regen_data_NewBackUp %>% 
  rename(plot = Gap_No) %>% 
  separate(plot, c("Gap", "Plot"))

SalvageLAI4 <- Regen_data_NewBackUp %>% 
  filter(Gap_type== "S") %>% 
  group_by(Gap, Plot, `LAI 4Ring`) %>% 
  summarise(count = n())
```

Salvage Scatter plots: Plotting scatterplot for number of trees per plot and LAI 4R per plot 
```{r}
SalvScatter4R <- SalvageLAI4 %>%  STotalSpeciesPlot %>%
PlotScatter4R <- plot(SalvageLAI4$`LAI 4Ring` ~ STotalSpeciesPlot$mean_plot)
```

Mean of natural species per plot to match LAI 4R data

```{r}
NTotalSpeciesPlot <- Natural_regen_summary_plots  %>%
  group_by(Gap, Plot) %>% 
  summarise(mean_plot=mean(TPHN)) 
View(NTotalSpeciesPlot)
```
Group LAI to match up natural regeneration data
```{r}
NaturalLAI4 <- Regen_data_NewBackUp %>% 
  filter(Gap_type== "N") %>% 
  group_by(Gap, Plot, `LAI 4Ring`) %>% 
  summarise(count = n())
```

Natural Scatter plots: Plotting scatterplot for number of trees per plot and LAI 4R per plot
```{r}
NatScatter4R <- NaturalLAI4 %>%  NTotalSpeciesPlot %>%
PlotNatScatter4R <- plot(NaturalLAI4$`LAI 4Ring` ~ NTotalSpeciesPlot$'mean_plot')
```

